---
title: "CMAC Brand in ggplot2 - Installation and Examples"
author: "Daniel Carpenter"
format: 
    gfm:
      toc: true
      toc-depth: 3
      number-sections: true
editor: visual
---

# `styles.cmac`

<!-- badges: start -->

<!-- badges: end -->

## Overview

> -   CMAC Brand in `ggplot2`
> -   Package that streamlines color palette usage, adapated from the [Tableau 10 Palette](https://www.tableau.com/blog/colors-upgrade-tableau-10-56782)
> -   See colorblind statistics for the [fill palette](https://davidmathlogic.com/colorblind/#%23BECDE0-%23FFD597-%23F6B7B4-%23BEE0D2-%23E4C6DC-%23BDDBE1-%23F7E5B3-%23DCDADA-%23FCC8DA-%23E5CFC5) and the [color (*line*) palette](https://davidmathlogic.com/colorblind/#%236388B4-%23E68900-%23EB4B43-%2354AC88-%23B07AA1-%233C9DAA-%23C5A952-%238A807E-%23D16D91-%23945430)
> -   ***PLEASE SEE THE PACKAGE [VIGNETTE]{.underline} FOR MORE DETAILED EXAMPLES***

<br>

## Installation of Package

You can install the development version of styles.cmac from [GitHub](https://github.com/) with:

```{r, eval=FALSE}
# Install devtools if not installed (for GitHub Package)
if (!require("devtools")) install.packages("devtools")

devtools::install_github("CN-OIMSF/styles.cmac")
```

## Viewing the Vignette

```{r}
# see what is available
browseVignettes('styles.cmac')

# To view it then run: 
vignette("intro-to-package", package = "styles.cmac")
```

## Examples

```{r}
library(styles.cmac)
```

### Create a basic plot structure

> Plot that does *NOT* include CMAC Styles

```{r}
library(ggplot2)

basePlot <- ggplot(data = mtcars,
               aes(x     = wt,
                   shape = as.factor(cyl),
                   color = as.factor(cyl),
                   y     = mpg
                   )
               ) +
          geom_point(size = 5)
basePlot # Display
```

### Use Multiple Colors

#### Default Color Palette

> Using CMAC Style

```{r}
basePlot +
  
  # Use the CMAC color palette
  scale_color_cmac() +
  
  # Get the CMAC ggplot theme
  theme_cmac()
```

#### Override Color with Fill

> Using CMAC Style

```{r}
basePlot +
  
  # Use the CMAC Fill palette (override)
  scale_color_cmac(overrideWithFill = TRUE) +  
  
  # Get the CMAC ggplot theme
  theme_cmac()
```

### Use only `1` Color

> Using CMAC Style

```{r}
basePlot + 
  
  # Use only a single color (note using line palette)
  geom_point(color = scale_cmac('blue', 'blue2'),
             size  = 5) +
  
  # Get the CMAC ggplot theme
  theme_cmac()
```

### How to use Times New Roman

> Use this setting for ***non-website*** reporting.

```{r}
basePlot +
  
  # Use the CMAC color palette
  scale_color_cmac() +
  
  # Get the CMAC ggplot theme
  theme_cmac(inTimesNewRoman = TRUE)
```

## Colors Deep Dive

### What Colors are Available?

-   Notice `displayNames = TRUE`, which show you what hex codes are associated with the color names.

-   Default is `displayNames = FALSE` for best functionality with plotting

#### Fill colors

```{r}
scale_cmac('fill', displayNames = TRUE)
```

#### Line colors

```{r}
scale_cmac('color', displayNames = TRUE)
```

#### Blue and Gray colors, like the background of slide decks

```{r}
scale_cmac('blue', displayNames = TRUE)
scale_cmac('gray', displayNames = TRUE)
```

#### Text, Grays, and Blues

-   Common colors that you might need

```{r}
scale_cmac('text') # Text (dark gray) 
scale_cmac('blue') # Grays that are in the brand
scale_cmac('gray') # Blues that are in the brand
```

### How to get `1` color?

```{r}
# Single color from the fill palette
scale_cmac('fill', 'red')

# Single color from the base palette
scale_cmac('gray', 'gray3')

# Single color from the base palette
scale_cmac('blue', 'blue2')
```

### How to get `>1` colors

```{r}
# Get the first 3 colors in the line palette
scale_cmac('color', 3)

# Get the last 3 colors in the fill palette
scale_cmac('color')[6:8]

# Or access specific colors all at once
scale_cmac('color', 'blue', 'orange', 'green', 'yellow')
```
