---
title: "A Guide to the styles.cmac Package"
output: 
  rmarkdown::html_vignette:
    toc: yes
vignette: >
  %\VignetteIndexEntry{A Guide to the styles.cmac Package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 5.5,
  fig.height = 4
)
```

## Overview

The `styles.cmac` package provides a suite of functions for color styling in `ggplot2`, with an emphasis on branding for CMAC. Notably, it includes a palette that is adapted from the Tableau 10 Palette. This package is designed to streamline the usage of color palettes, with inspiration drawn from the Tableau 10 Palette.

For a deeper dive into the package's colorblind-friendly design, visit the fill palette and color (*line*) palette URLs provided below. This gives an overview of what the palette looks like as a whole.

-   See colorblind statistics for the [fill palette](https://davidmathlogic.com/colorblind/#%23B4D0F4-%23FFD39F-%23FFB5B3-%23B4DFCE-%23CDC9E0-%23F2D4A0-%23C3C6D2-%23E6D7D7) and the [color (*line*) palette](https://davidmathlogic.com/colorblind/#%237996B9-%23DBA657-%23C86D6A-%2383AF9E-%23928EA4-%23C2A570-%23828591-%23BAABAB)

-   [*Reference*: Original Tableau 10 Palette](https://www.tableau.com/blog/colors-upgrade-tableau-10-56782)

## Installation

If you want to use the latest version of `styles.cmac` from GitHub, ensure that you have the devtools package installed, and then use devtools::install_github() to download and install `styles.cmac.`

```{r, eval=FALSE}
# If devtools is not installed
if (!require("devtools"))   install.packages("devtools")

# Install styles.cmac from GitHub
devtools::install_github("CN-CMAC/styles.cmac")
```

## Examples

### Setup

First, we load the `styles.cmac` package.

```{r}
library(styles.cmac)
```

### Adjusting `color` Aesthetic

Let's dive into some examples of how to use styles.cmac in conjunction with `ggplot2.` First, we will create a basic plot using `ggplot2`, without any CMAC styling.

```{r}
library(ggplot2)

# Create a basic plot structure
basePlot <- ggplot(data = mtcars,
                   aes(x     = wt,
                       shape = as.factor(cyl),
                       color = as.factor(cyl),
                       y     = mpg
                   )
) +
  geom_point(size = 5)
basePlot # Display
```

Now, let's look at how we can use the CMAC color palette and theme with our base plot.

```{r}
basePlot +
  
  # Use the CMAC color palette
  scale_color_cmac() +
  
  # Get the CMAC ggplot theme
  theme_cmac()
```

This plot now properly leverages the CMAC brand.

### Other Options with `color` Aesthetic

Here are some other useful examples that demonstrate some of the function's flexibility.

```{r}
# Use the accent color palette instead of the fill
# This would be beneficial if you need to use the fill colors instead of the accent colors.
basePlot + scale_color_cmac(overrideWithFill = TRUE) + theme_cmac()

# Add the fill color and turn off legend with optional parameter
# You can use any parameter supplied through ggplot2::scale_fill_manual()
basePlot + scale_color_cmac(guide = 'none') + theme_cmac()

# You could also change the theme through the optional parameters
# Add the theme with additional optional theme parameters
basePlot + scale_color_cmac() + theme_cmac(legend.position = 'right')
```

Additional parameter called `borderMode` controls the theme borders. [Inspiration for facets came from 1.2.2 in this book.](https://socviz.co/lookatdata.html)

```{r}
# One other option is to remove the border when faceting
# Note this example is only for demonstrating the borders. Faceting in
# this context is another issue in itself. 
basePlot + scale_color_cmac() + theme_cmac(borderMode = 'facet') +
  facet_grid(~as.factor(cyl))

# Note you can also use 'borders' to add all border back.
basePlot + scale_color_cmac() + theme_cmac(borderMode = 'borders')
```

### Use only `1 color`

Most of the time you only need one color. Here is an example of how you might get that.

```{r}
basePlot + 
  
  # Use only a single color (note using line palette)
  geom_point(color = scale_cmac('blue', 'blue2'),
             size  = 5) +
  
  # Get the CMAC ggplot theme
  theme_cmac()
```

### Adjusting `fill` Aesthetic

Here are some examples of how you could use the fill colors in plots.

Create a Base plot

```{r}
# Create basic ggplot object
ggplotObject <- ggplot(mtcars, aes(y = mpg, x = as.factor(cyl), 
                                   fill = as.factor(cyl), color = as.factor(cyl))) +
  
  # Note that gray border used for demonstration. Please use scale_color_cmac()
  # as much as possible
  geom_boxplot(size = 0.8)
```

Here is the main combination that you might use.

```{r}
# Add the fill and color
ggplotObject + scale_fill_cmac() + scale_color_cmac()
```

Below shows more options that demonstates some of the function's flexibility.

```{r}
# Create basic ggplot object
ggplotObject2 <- ggplot(mtcars, aes(y = mpg, x = as.factor(cyl), fill = as.factor(cyl))) +
  
  # Note that gray border used for demonstration. Please use scale_color_cmac()
  # as much as possible
  geom_boxplot(size = 0.8, color = scale_cmac('gray', 'gray3'))

# Add the accent color
ggplotObject2 + scale_fill_cmac(overrideWithAccent = TRUE)

# Add the fill color and turn off legend with optional parameter
ggplotObject2 + scale_fill_cmac(guide = 'none')
```

### How to use *Times New Roman*

> Use this setting for ***non-website*** reporting, i.e., PDF distribution.

#### Non-Label Font Control

```{r}
basePlot +
  
  # Use the CMAC color palette
  scale_color_cmac() +
  
  # Get the CMAC ggplot theme
  theme_cmac(inTimesNewRoman = TRUE)
```

#### Label Font Control

```{r}
basePlot + 
  
  geom_text(aes(label = paste(mpg)),
            vjust = 1.2, hjust = -0.2,
            size = 3.5,
            
            # ** IMPORTANT: theme() does not control text labels
            # Need to set family to this font!
            family = styles.cmac::getTimesNewRomanFamily()
            ) + 
  
  # This controls all non-label fonts
  theme_cmac(inTimesNewRoman = TRUE)
```

### Darken Text when Over Filled Elements

```{r}
ggplot(mtcars, aes(y = mpg, x = wt, color = as.factor(cyl))) +
  geom_point(aes(fill = as.factor(cyl)),
             size = 10,
             pch = 21,
             color = 'transparent',
             alpha = 0.9
  ) +
  geom_text(aes(label = round(mpg, 0))) +
  scale_fill_cmac() +

  # KEY - darken the text so that it is easier to view
  # Over fill
  scale_color_cmac(darkenPaletteForTextGeoms = TRUE) +
  theme_cmac()
```

### Offset or Reverse Order of Colors

```{r}
# Offset the colors by 1
basePlot + scale_color_cmac(colorOffset = 1)

# reverse the order of the palette
basePlot + scale_color_cmac(reverseOrder = TRUE)
```

### Override the palette with a color blind palette

```{r}
# Use color blind friendly palette (works with fill too)
basePlot + scale_color_cmac(useColorBlindPalette = TRUE)

# Change the palette (can use cols4all::c4a_palettes() to try others)
# Also, can demo others in GUI using cols4all::c4a()
basePlot + scale_color_cmac(useColorBlindPalette = TRUE, 
                            colorBlindPaletteName = 'color_blind')
```

## Numeric Formats

> Mainly for quick financial axis formatting

```{r}
kDollarsFormat(1000,          scaleUnit = 'K')
kDollarsFormat(1000000,       scaleUnit = 'M')
kDollarsFormat(1000000,       scaleUnit = 'MM')
kDollarsFormat(1000000000,    scaleUnit = 'B')
kDollarsFormat(1500000000000, scaleUnit = 'T')
kDollarsFormat(1000000,       scaleUnit = 'M', useDollarSign = FALSE)
```

## Colors Deep Dive

This section take a closer look at the colors that are available to us in the `styles.cmac` package.

### What Colors are Available?

Notice `displayNames = TRUE`, which show you what hex codes are associated with the color names.

Default is `displayNames = FALSE` for best functionality with plotting.

```{r}
# Fill colors
scale_cmac('fill', displayNames = TRUE)

# Line colors
scale_cmac('color', displayNames = TRUE)

# Blue and Gray colors, like the background of slide decks
scale_cmac('blue', displayNames = TRUE)
scale_cmac('gray', displayNames = TRUE)

# Text, Grays, and Blues
scale_cmac('text') # Text (dark gray) 
scale_cmac('blue') # Grays that are in the brand
scale_cmac('gray') # Blues that are in the brand
```

### How to get `1 color`?

```{r}
# Single color from the fill palette
scale_cmac('fill', 'red')

# Single color from the base palette
scale_cmac('gray', 'gray3')

# Single color from the base palette
scale_cmac('blue', 'blue2')
```

### How to get `> 1 colors`

```{r}
# Get the first 3 colors in the line palette
scale_cmac('color', 3)

# Get the last 3 colors in the fill palette
scale_cmac('color')[6:8]

# Or access specific colors all at once
scale_cmac('color', 'blue', 'orange', 'green', 'yellow')
```
