---
title: "A Guide to the styles.cmac Package"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{A Guide to the styles.cmac Package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Setup

First, we load the `styles.cmac` package. This package is designed to streamline the usage of color palettes, with inspiration drawn from the Tableau 10 Palette. 

```{r}
library(styles.cmac)
```

## Overview

The `styles.cmac` package provides a suite of functions for color styling in `ggplot2`, with an emphasis on branding for CMAC. Notably, it includes a palette that is adapted from the Tableau 10 Palette. For a deeper dive into the package's colorblind-friendly design, visit the fill palette and color (*line*) palette URLs provided in the original document.

## Installation

If you want to use the latest version of `styles.cmac` from GitHub, ensure that you have the devtools package installed, and then use devtools::install_github() to download and install `styles.cmac.`

```{r, eval=FALSE}
# If devtools is not installed
if (!require("devtools"))   install.packages("devtools")

# Install styles.cmac from GitHub
devtools::install_github("CN-OIMSF/styles.cmac")
```

## Examples

Let's dive into some examples of how to use styles.cmac in conjunction with `ggplot2.` First, we will create a basic plot using `ggplot2`, without any CMAC styling.

```{r}
library(ggplot2)

# Create a basic plot structure
basePlot <- ggplot(data = mtcars,
                   aes(x     = wt,
                       shape = as.factor(cyl),
                       color = as.factor(cyl),
                       y     = mpg
                       )
                   ) +
            geom_point(size = 5)
basePlot # Display
```

Now, let's look at how we can use the CMAC color palette and theme with our base plot.

```{r}
basePlot +
  
  # Use the CMAC color palette
  scale_color_cmac() +
  
  # Get the CMAC ggplot theme
  theme_cmac()
```

This plot is looking much better already!

### Override Color with Fill

```{r}
basePlot +
  
  # Use the CMAC Fill palette (override)
  scale_color_cmac(overrideWithFill = TRUE) +  
  
  # Get the CMAC ggplot theme
  theme_cmac()
```

### Use only `1` Color

```{r}
basePlot + 
  
  # Use only a single color (note using line palette)
  geom_point(color = scale_cmac('blue', 'blue2'),
             size  = 5) +
  
  # Get the CMAC ggplot theme
  theme_cmac()
```

## How to use Times New Roman

> Use this setting for ***non-website*** reporting.

```{r}
basePlot +
  
  # Use the CMAC color palette
  scale_color_cmac() +
  
  # Get the CMAC ggplot theme
  theme_cmac(inTimesNewRoman = TRUE)
```

## Colors Deep Dive

Let's take a closer look at the colors that are available to us in the `styles.cmac` package. 

### What Colors are Available?

Notice `displayNames = TRUE`, which show you what hex codes are associated with the color names.

Default is `displayNames = FALSE` for best functionality with plotting.

```{r}
# Fill colors
scale_cmac('fill', displayNames = TRUE)

# Line colors
scale_cmac('color', displayNames = TRUE)

# Blue and Gray colors, like the background of slide decks
scale_cmac('blue', displayNames = TRUE)
scale_cmac('gray', displayNames = TRUE)

# Text, Grays, and Blues
scale_cmac('text') # Text (dark gray) 
scale_cmac('blue') # Grays that are in the brand
scale_cmac('gray') # Blues that are in the brand
```

### How to get `1` color?

```{r}
# Single color from the fill palette
scale_cmac('fill', 'red')

# Single color from the base palette
scale_cmac('gray', 'gray3')

# Single color from the base palette
scale_cmac('blue', 'blue2')
```

### How to get `>1` colors

```{r}
# Get the first 3 colors in the line palette
scale_cmac('color', 3)

# Get the last 3 colors in the fill palette
scale_cmac('color')[6:8]

# Or access specific colors all at once
scale_cmac('color', 'blue', 'orange', 'green', 'yellow')
```
